The \project{unRAVE} catalog

### 2016-08-18 article/unrave.tex
Some filler words for internal validation tests

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge.
The $I$ band was used because it overlaps with the wavelength range 
that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

@@

\section{Validation}
\label{sec:validation}

% Chi-sq of the sample.
% Identify outliers.
% Convex hull? Distance to plane?

% Stars with morphological classification

% Internal validation:
% --> repeat visits: are our formal errors reasonable?
% --> bootstrap resampling of the training set?

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> APOGEE / EPIC (whichever was not used)
% --> Literature studies presented in Kordopatis+ 2014

% Astrophysical validation:
% --> Identify open clusters and look at their metallicity spreads
% --> Identify globular clusters and look to see whether we capture "expected" abundance trends/spreads that are seen elsewhere in the literature
% --> Residuals as a function of galactic position? (DIBS)

@@

% TODO: Replace ZENODO-URL with a data url
Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  
This document was compiled from revision hash \texttt{\githash} in that repository.
Inferred labels and covariance matrices are available electronically with 
associated metadata at \texttt{\url{ZENODO-URL}}.  Please note that it is a condition
of using these results that the fifth \rave\ data release by \citet{Kunder_2016} 
must (also) be cited, as the work presented here would not have been possible 
without the tireless efforts of the entire \rave\ collaboration, past and present.

>>>>

The target selection for \rave\ is based on the $I$-band apparent magnitude,
$9 < I < 12$, with a weak $J - K_s > 0.5$ cut near the disk and bulge\footnote{
However, see the works of \citet{Kunder_2016} and \citet{Kordopatis_2013} for 
subtleties of the selection function}.
The $I$ band was used because it overlaps with the wavelength range 
that \rave\ operates in:  8410--8795~\AA.  This setup includes the \ion{Ca}{2} 
near-infrared triplet.  These stellar absorption lines are dominated by pressure 
broadening, which makes them strong and therefore visible even in very low S/N data
or metal-poor stars.  This spectral region is also well-studied.  It is one of the
key setups used for the ground-based high-resolution \project{Gaia-ESO} survey
\citep{Gilmore_2012,Randich_2013}, and the \rave\ resolution and wavelength
coverage is comparable to the Radial Velocity Spectrometer on board the \project{Gaia}
space telescope \citep{Recio-Blanco_2016}.

@@

\section{Validation experiments}
\label{sec:validation}

\subsection{Internal validation}
We first conducted leave-one-out cross-validation to validate our results.
For each training set star we removed that star from the training set and
then re-trained the model.  Using that re-trained model, we then treated
the excluded star as a test star and compared our inferred labels to the
test set labels.  We show the results of this experiment in Figure
\ref{fig:loocv-comparison}.  The bias and root mean square (RMS) error in 
$\teff$ is XX~K and XX~K, and in $\logg$ is X.XX~dex and X.XX~dex,
respectively.  The bias in abundance labels is negligible --- between
X.XX to X.XX --- and the RMS varies from X.XX ([XX/H]) to X.XX ([XX/H]).

The \rave\ survey performed repeat observations for 23,288 stars  with time 
intervals ranging from a few hours to up to four years.  This timing was 
constructed to be quasi-logarithmic such that spectroscopic binaries could
be optimally identified. Most of the stars that were observed multiple times
were only observed twice, with thirteen visits being the maximum number 
of observations for any target.  These repeat observations allow us to 
quantify the level of (in)correctness in our formal errors.  For every star
with multiple visits we constructed a high S/N stacked spectrum for that
star, weighted by the inverse variances in the individual visit spectra,
which then served as the basis of comparison to all individual visits.  
Figure \ref{fig:formal-errors-comparison} shows the difference in labels 
between the comparison spectrum and a repeat visit, normalized by their 
formal errors summed in quadrature (e.g., $\Delta\logg/\sqrt{\sigma_{\logg,stacked}^2 + \sigma_{\logg,visit}^2}$).
If our measurements were unbiased by S/N and the errors were representative, 
these values should be normally distributed with a zero mean and unit variance.
\stub{Are they? Should we inflate our formal errors?}

% Internal validation:
% --> bootstrap resampling of the training set?

% Identify outliers:
% --> Chi-sq of the sample.
% --> Convex hull? Distance to plane?
% --> Stars with morphological classification
% --> Anything else?

\subsection{External validation}

% External validation: 
% --> RAVE Data Release 4
% --> Gaia-ESO Survey
% --> APOGEE / EPIC (whichever was not used)
% --> Literature studies presented in Kordopatis+ 2014
% --> Gold standards (incl. exoplanet host star studies): Reddy (2003, 2006), Bensby (2014), Valenti & Fischer (2005)

\subsection{Astrophysical validation}
% Astrophysical validation:
% --> Identify globular clusters and look to see whether we capture "expected" abundance trends/spreads that are seen elsewhere in the literature
% --> Residuals as a function of galactic position? (DIBS)

@@

% TODO: Replace ZENODO-URL with a data url
Source code for this project is available at \texttt{\url{https://www.github.com/AnnieJumpCannon/RAVE}}.  
This document was compiled from revision hash \texttt{\githash} in that repository.
Inferred labels and covariance matrices are available electronically with 
associated metadata at \texttt{\url{ZENODO-URL}} \citep{DATA_REPOSITORY}.  
Please note that it is a condition of using these results that the fifth \rave\ 
data release by \citet{Kunder_2016} must (also) be cited, as the work presented 
here would not have been possible without the tireless efforts of the entire 
\rave\ collaboration, past and present.
